---
title: "Final Exam: Part 1"
output: 
  html_notebook: 
    highlight: pygments
    theme: flatly
---

## Overview
**Forecasting Interest Rate Hikes by the U.S. Federal Reserve**

The federal funds rate is the key interest rate that the U.S. Federal Reserve uses to influence economic growth. The Federal Open Market Committee meets regularly to decide whether to increase, decrease, or maintain the target interest rate. Their choice has important ramifications that cascade through the economy, so the announcement of the interest rates is eagerly awaited each month.

In this problem, we will use analytics to try to predict when the Federal Reserve will raise interest rates. We will look at monthly economic and political data dating back to the mid-1960's. In this analysis, the dependent variable will be the binary outcome variable **RaisedFedFunds**, which takes value 1 if the federal funds rate was increased that month and 0 if it was lowered or stayed the same. For each month, the file federalFundsRate.csv contains the following independent variables:

- **Date**: The date the change was announced.
- **Chairman**: The name of the Federal Reserve Chairman at the time the change was announced.
- **PreviousRate**: The federal funds rate in the prior month.
- **Streak**: The current streak of raising or not raising the rate, e.g. +8 indicates the rate has been increased 8 months in a row, whereas -3 indicates the rate has been lowered or stayed the same for 3 months in a row.
- **GDP**: The U.S. Gross Domestic Product, in Billions of Chained 2009 US Dollars.
- **Unemployment**: The unemployment rate in the U.S.
- **CPI**: The Consumer Price Index, an indicator of inflation, in the U.S.
- **HomeownershipRate**: The rate of homeownership in the U.S.
- **DebtAsPctGDP**: The U.S. national debt as a percentage of GDP
- **DemocraticPres**: Whether the sitting U.S. President is a Democrat (DemocraticPres=1) or a Republican (DemocraticPres=0)
- **MonthsUntilElection**: The number of remaining months until the next U.S. presidential election.

## Solutions
### Initialization
```{r, results='hide' ,warning=FALSE}
library(dplyr)
library(magrittr)
```

### Problem 1 - Loading the Data
Use the `read.csv()` function to load the contents of `federalFundsRate.csv` into a data frame called `fedFunds`, using `stringsAsFactors=FALSE`.
```{r}

```

#### Question:
What proportion of months did the Fed raise the interest rate?

#### Answer:
```{r}

```

### Problem 2 - The Longest-Serving Fed Chair

#### Question:
Which Federal Reserve Chair has presided over the most interest rate decisions?

#### Answer:
```{r}

```

### Problem 3 - Converting Variables to Factors
Convert the following variables to factors using the as.factor function:
- Chairman
- DemocraticPres
- RaisedFedFunds

```{r}

```

#### Question:
Which of the following methods requires the dependent variable be stored as a factor variable when training a model for classification?
Logistic regression (glm)
CART (rpart)
Random forest (randomForest)

#### Answer:
**Answer**

### Problem 4 - Splitting into a Training and Testing Set

Obtain a random training/testing set split with:
```{r}
set.seed(201)
library(caTools)
spl <- sample.split(fedFunds$RaisedFedFunds, 0.7)
```

Split months into a training data frame called `training` using the observations for which `spl == TRUE` and a testing data frame called `testing` using the observations for which `spl == FALSE`.

#### Question:
Why do we use the `sample.split()` function to split into a training and testing set?
- It is the most convenient way to randomly split the data
- It balances the independent variables between the training and testing sets
- It balances the dependent variable between the training and testing sets

#### Answer:


### Problem 5 - Training a Logistic Regression Model
Train a logistic regression model using independent variables `PreviousRate`, `Streak`, `Unemployment`, `HomeownershipRate`, `DemocraticPres`, and `MonthsUntilElection`, using the training set to obtain the model.

#### Question:
Which of the following characteristics is the most statistically significant associated with an increased chance of the federal funds rate being raised?
- A higher federal funds rate in the previous month.  A longer consecutive streak of months in which the federal funds rate was raised.  A higher unemployment rate.  A higher rate of homeownership.  The current president being a Democrat.  A longer amount of time until the next presidential election.

#### Answer:
```{r}

```

### Problem 6 - Predicting Using a Logistic Regression Model
Imagine you are an analyst at a bank and your manager has asked you to predict whether the federal funds rate will be raised next month. You know that the rate has been lowered for 3 straight months (Streak = -3) and that the previous month's rate was 1.7%. The unemployment rate is 5.1% and the homeownership rate is 65.3%. The current U.S. president is a Republican and the next election will be held in 18 months.

#### Question:
According to the logistic regression model you built in Problem 5, what is the predicted probability that the interest rate will be raised?

#### Answer:
```{r}

```

### Problem 7 - Interpreting Model Coefficients

#### Question:
What is the meaning of the coefficient labeled `DemocraticPres1` in the logistic regression summary output?
- When the president is Democratic, the odds of the federal funds rate increasing are 34.8% higher than in an otherise identical month (i.e. identical among the variables in the model).  When the president is Democratic, the odds of the federal funds rate increasing are 34.8% higher than in an average month in the dataset.  When the president is Democratic, the odds of the federal funds rate increasing are 41.6% higher than in an otherise identical month (i.e. identical among the variables in the model).  When the president is Democratic, the odds of the federal funds rate increasing are 41.6% higher than in an average month in the dataset.

#### Answer:
```{r}

```

### Problem 8 - Obtaining Test Set Predictions
Using your logistic regression model, obtain predictions on the test set. Then, using a probability threshold of 0.5, create a confusion matrix for the test set.

We would like to compare the predictions obtained by the logistic regression model and those obtained by a naive baseline model. Remember that the naive baseline model we use in this class always predicts the most frequent outcome in the training set for all observations in the test set.

#### Question:
What is the number of test set observations where the prediction from the logistic regression model is different than the prediction from the baseline model?

#### Answer:
```{r}

```

### Problem 9 - Computing Test-Set AUC

#### Question:
What is the test-set AUC of the logistic regression model?

#### Answer:
```{r}

```

### Problem 10 - Interpreting AUC

#### Question:
What is the meaning of the AUC?
- The proportion of the time the model can differentiate between a randomly selected month during which the federal funds were raised and a randomly selected month during which the federal funds were not raised
- The proportion of the time the model correctly identifies whether or not federal funds were raised

#### Answer:
```{r}

```

### Problem 11 - ROC Curves

#### Question:
Which logistic regression threshold is associated with the upper-right corner of the ROC plot (true positive rate 1 and false positive rate 1)?
- 0
- 0.5
- 1

#### Answer:
```{r}

```

### Problem 12 - ROC Curves
Plot the colorized ROC curve for the logistic regression model's performance on the test set.

#### Question:
At roughly which logistic regression cutoff does the model achieve a true positive rate of 85% and a false positive rate of 60%?
- 0  0.16  0.37  0.52  0.66  0.87
 
#### Answer:
```{r}

```

### Problem 13 - Cross-Validation to Select Parameters

#### Question:
Which of the following best describes how 10-fold cross-validation works when selecting between 2 different parameter values?
- 2 models are trained on subsets of the training set and evaluated on the testing set  10 models are trained on subsets of the training set and evaluated on the testing set  20 models are trained on subsets of the training set and evaluated on the testing set  2 models are trained on subsets of the training set and evaluated on a portion of the training set  10 models are trained on subsets of the training set and evaluated on a portion of the training set  20 models are trained on subsets of the training set and evaluated on a portion of the training set

#### Answer:
```{r}

```

### Problem 14 - Cross-Validation for a CART Model
Set the random seed to 201 (even though you have already done so earlier in the problem). Then use the caret package and the train function to perform 10-fold cross validation with the training data set to select the best cp value for a CART model that predicts the dependent variable `RaisedFedFunds` using the independent variables `PreviousRate`, `Streak`, `Unemployment`, `HomeownershipRate`, `DemocraticPres`, and `MonthsUntilElection`. Select the `cp` value from a grid consisting of the 50 values 0.001, 0.002, ..., 0.05.

```{r}

```


#### Question:
What `cp` value maximizes the cross-validation accuracy?

#### Answer:
```{r}

```

### Problem 15 - Train CART Model
Build and plot the CART model trained with the parameter identified in Problem 14, again predicting the dependent variable using `PreviousRate`, `Streak`, `Unemployment`, `HomeownershipRate`, `DemocraticPres`, and `MonthsUntilElection`.

#### Question:
What variable is used as the first (upper-most) split in the tree?
- PreviousRate  Streak  Unemployment  HomeownershipRate  DemocraticPres  MonthsUntilElection

#### Answer:
```{r}

```

### Problem 16 - Predicting Using a CART Model
Remember: The rate has been lowered for 3 straight months (Streak = -3). The previous month's rate was 1.7%. The unemployment rate is 5.1%. The homeownership rate is 65.3%. The current U.S. president is a Republican and the next election will be held in 18 months.

#### Question:
If you were to use the CART model you created in Problem 15 to answer the question asked of the analyst in Problem 6, what would you predict for next month?
- The Fed will raise the federal funds rate.  The Fed will not raise the federal funds rate.

#### Answer:
```{r}

```

### Problem 17 - Test-Set Accuracy for CART Model
Using the CART model you created in Problem 15, obtain predictions on the test set (using the parameter type="class" with the predict function). Then, create a confusion matrix for the test set.

#### Question:
What is the accuracy of your CART model?

#### Answer:
```{r}

```

