---
title: "Final Exam: Part 1"
output: html_notebook
---
## Overview
FORECASTING INTEREST RATE HIKES BY THE U.S. FEDERAL RESERVE

The federal funds rate is the key interest rate that the U.S. Federal Reserve uses to influence economic growth. The Federal Open Market Committee meets regularly to decide whether to increase, decrease, or maintain the target interest rate. Their choice has important ramifications that cascade through the economy, so the announcement of the interest rates is eagerly awaited each month.

In this problem, we will use analytics to try to predict when the Federal Reserve will raise interest rates. We will look at monthly economic and political data dating back to the mid-1960's. In this analysis, the dependent variable will be the binary outcome variable RaisedFedFunds, which takes value 1 if the federal funds rate was increased that month and 0 if it was lowered or stayed the same. For each month, the file federalFundsRate.csv contains the following independent variables:

- Date: The date the change was announced.
- Chairman: The name of the Federal Reserve Chairman at the time the change was announced.
- PreviousRate: The federal funds rate in the prior month.
- Streak: The current streak of raising or not raising the rate, e.g. +8 indicates the rate has been increased 8 months in a row, whereas -3 indicates the rate has been lowered or stayed the same for 3 months in a row.
- GDP: The U.S. Gross Domestic Product, in Billions of Chained 2009 US Dollars.
- Unemployment: The unemployment rate in the U.S.
- CPI: The Consumer Price Index, an indicator of inflation, in the U.S.
- HomeownershipRate: The rate of homeownership in the U.S.
- DebtAsPctGDP: The U.S. national debt as a percentage of GDP
- DemocraticPres: Whether the sitting U.S. President is a Democrat (DemocraticPres=1) or a Republican (DemocraticPres=0)
- MonthsUntilElection: The number of remaining months until the next U.S. presidential election.

## Solutions
### Initialization
```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(magrittr)
```

### Problem 1 - Loading the Data
Use the read.csv function to load the contents of federalFundsRate.csv into a data frame called fedFunds, using stringsAsFactors=FALSE.
```{r}

```

#### Question:
What proportion of months did the Fed raise the interest rate?

#### Answer:
```{r}

```

### Problem 2 - The Longest-Serving Fed Chair

#### Question:
Which Federal Reserve Chair has presided over the most interest rate decisions?

#### Answer:
```{r}

```

### Problem 3 - Converting Variables to Factors
Convert the following variables to factors using the as.factor function:

- Chairman
- DemocraticPres
- RaisedFedFunds

```{r}

```

#### Question:
Which of the following methods requires the dependent variable be stored as a factor variable when training a model for classification?
Logistic regression (glm)
CART (rpart)
Random forest (randomForest)

#### Answer:
**Answer**

### Problem 4 - Splitting into a Training and Testing Set

Obtain a random training/testing set split with:

set.seed(201)

library(caTools)

spl = sample.split(fedFunds$RaisedFedFunds, 0.7)

Split months into a training data frame called "training" using the observations for which spl is TRUE and a testing data frame called "testing" using the observations for which spl is FALSE.

#### Question:
Why do we use the sample.split() function to split into a training and testing set?
- It is the most convenient way to randomly split the data
- It balances the independent variables between the training and testing sets
- It balances the dependent variable between the training and testing sets

#### Answer:


### Problem 5 - Training a Logistic Regression Model

#### Question:


#### Answer:
```{r}

```

### Problem 6 - Predicting Using a Logistic Regression Model

#### Question:


#### Answer:
```{r}

```

### Problem 7 - Interpreting Model Coefficients

#### Question:


#### Answer:
```{r}

```

### Problem 8 - Obtaining Test Set Predictions

#### Question:


#### Answer:
```{r}

```

### Problem 9 - Computing Test-Set AUC

#### Question:


#### Answer:
```{r}

```

### Problem 10 - Interpreting AUC

#### Question:


#### Answer:
```{r}

```

### Problem 11 - ROC Curves

#### Question:


#### Answer:
```{r}

```

### Problem 12 - ROC Curves

#### Question:


#### Answer:
```{r}

```

### Problem 13 - Cross-Validation to Select Parameters

#### Question:


#### Answer:
```{r}

```

### Problem 14 - Cross-Validation for a CART Model

#### Question:


#### Answer:
```{r}

```

### Problem 15 - Train CART Model

#### Question:


#### Answer:
```{r}

```



### Problem 16 - Predicting Using a CART Model
#### Question:


#### Answer:
```{r}

```



### Problem 17 - Test-Set Accuracy for CART Model

#### Question:


#### Answer:
```{r}

```

